week getbackontrack

### 2024-11-18 clarifyRQsandDesign.Rmd
visualtextdarkmodecorrections

>>>>

# 17.11.24:

1.  **RQ1**: To what extent does gender homophily occur across all co-publications of the staff in the Sociology and Political Science Departments at Radboud University?\
    *This question focuses on descriptive analysis at the network level, examining overall gender clustering patterns within and between departments.*

2.  **RQ2**: To what extent do the two departments differ in their gender homophily?\
    *This question compares the strength of gender homophily between the two departments, with a focus on dyadic-level interactions while controlling for structural network effects.*

3.  **RQ3**: How does gender homophily in collaborations differ depending on scholars’ prestige (e.g., publication in high-impact journals)?\
    *This question investigates the role of prestige in moderating gender homophily at the ego level.*

### 2024-11-18 descriptivestry.Rmd
visualtextdarkmodecorrections

### Plotting: h-index -ggraph
Credits to Hannah Groennou for the basis of the code of these graphs. 
#### with isolates ?
```{r}

@@

palette.colors(palette = "-palette name here-")
# [1] "#000000" "#E69F00" "#56B4E9" "#009E73" "#F0E442" "#0072B2"
# [7] "#D55E00" "#CC79A7" "#999999"
```

>>>>

### Plotting: h-index -ggraph
Credits to Hannah Groennou for the basis of the code of these graphs.

(test line)

#### with isolates ?
```{r}

@@

palette.colors(palette = "Alphabet")
#R3
"#000000" "#FF0000" "#00CD00" "#0000FF" "#00FFFF" "#FF00FF" "#FFFF00" "#BEBEBE"

#R4
"#000000" "#DF536B" "#61D04F" "#2297E6" "#28E2E5" "#CD0BBC" "#F5C710" "#9E9E9E"

#ggplot2
"#000000" "#F8766D" "#00BA38" "#619CFF" "#00BFC4" "#F564E3" "#B79F00" "#9E9E9E"

#Okabe Ito 
"#000000" "#E69F00" "#56B4E9" "#009E73" "#F0E442" "#0072B2" "#D55E00" "#CC79A7" "#999999"

# Accent
"#7FC97F" "#BEAED4" "#FDC086" "#FFFF99" "#386CB0" "#F0027F" "#BF5B17" "#666666"

#Paired
"#A6CEE3" "#1F78B4" "#B2DF8A" "#33A02C" "#FB9A99" "#E31A1C" "#FDBF6F" "#FF7F00"
"#CAB2D6" "#6A3D9A" "#FFFF99" "#B15928"

#Pastel 1
"#FBB4AE" "#B3CDE3" "#CCEBC5" "#DECBE4" "#FED9A6" "#FFFFCC" "#E5D8BD" "#FDDAEC" "#F2F2F2"

#Pastel 2
"#B3E2CD" "#FDCDAC" "#CBD5E8" "#F4CAE4" "#E6F5C9" "#FFF2AE" "#F1E2CC" "#CCCCCC"

#Set 1
"#E41A1C" "#377EB8" "#4DAF4A" "#984EA3" "#FF7F00" "#FFFF33" "#A65628" "#F781BF""#999999"

#Set 2
"#66C2A5" "#FC8D62" "#8DA0CB" "#E78AC3" "#A6D854" "#FFD92F" "#E5C494" "#B3B3B3"

#Set 3
"#8DD3C7" "#FFFFB3" "#BEBADA" "#FB8072" "#80B1D3" "#FDB462" "#B3DE69" "#FCCDE5"
"#D9D9D9" "#BC80BD" "#CCEBC5" "#FFED6F"

# Tableau 10
"#4E79A7" "#F28E2B" "#E15759" "#76B7B2" "#59A14F" "#EDC948" "#B07AA1" "#FF9DA7"
"#9C755F" "#BAB0AC"

#Classic Tableau
"#1F77B4" "#FF7F0E" "#2CA02C" "#D62728" "#9467BD" "#8C564B" "#E377C2" "#7F7F7F"
"#BCBD22" "#17BECF"

# Polychrome 36:
"#5A5156" "#E4E1E3" "#F6222E" "#FE00FA" "#16FF32" "#3283FE" "#FEAF16" "#B00068"
"#1CFFCE" "#90AD1C" "#2ED9FF" "#DEA0FD" "#AA0DFE" "#F8A19F" "#325A9B" "#C4451C"
"#1C8356" "#85660D" "#B10DA1" "#FBE426" "#1CBE4F" "#FA0087" "#FC1CBF" "#F7E1A0"
"#C075A6" "#782AB6" "#AAF400" "#BDCDFF" "#822E1C" "#B5EFB5" "#7ED7D1" "#1C7F93"
"#D85FF7" "#683B79" "#66B0FF" "#3B00FB"

#Alphabet
"#AA0DFE" "#3283FE" "#85660D" "#782AB6" "#565656" "#1C8356" "#16FF32" "#F7E1A0"
"#E2E2E2" "#1CBE4F" "#C4451C" "#DEA0FD" "#FE00FA" "#325A9B" "#FEAF16" "#F8A19F"
"#90AD1C" "#F6222E" "#1CFFCE" "#2ED9FF" "#B10DA1" "#C075A6" "#FC1CBF" "#B00068"
"#FBE426" "#FA0087"

### 2024-11-18 draftdata.Rmd
visualtextdarkmodecorrections

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

@@

```{r klippy, echo=FALSE, include=FALSE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

@@

[**CHECK AGAIN BEFORE SUBMITTING**]:\
A total of four models, including a null model, were estimated. The null model excludes structural effects, focusing solely on reciprocity and density. The first model incorporates all structural network effects. Model 2 builds on this by adding the ego-level effect of gender. In Model 3, the ego-level control variable for h-index is included. \
**[NEEDED?]** Finally, Model 4 is similar to Model 3 but replaces the quadratic effect of indegree popularity with a simple structural effect of indegree popularity.

@@

```{=html}
<!-- -->
```
-   We developed the Stata command h_index and R package hindex.

@@

-   Csárdi G, Nepusz T, Traag V, Horvát Sz, Zanini F, Noom D, Müller K (2024). \_igraph: Network Analysis and Visualization in R\_. doi:10.5281/zenodo.7682609 \<https://doi.org/10.5281/zenodo.7682609\>, R package version 2.1.1.9005, \<https://CRAN.R-project.org/package=igraph\>.

@@

-   Pedersen T (2024). *ggraph: An Implementation of Grammar of Graphics for Graphs and Networks*. R package version 2.2.1.9000, https://github.com/thomasp85/ggraph, [https://ggraph.data-imaginist.com](https://ggraph.data-imaginist.com/).

>>>>

```{=html}
<style>
body {
text-align: left;
line-height: 1.5;
}

h1, .h1, h2, .h2, h3, .h3 {
  margin-top: 24px;
  line-height: 1;
}

</style>
```
```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

@@

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

@@

@@

<!-- -->

-   We developed the Stata command h_index and R package hindex.

@@

-   Csárdi G, Nepusz T, Traag V, Horvát Sz, Zanini F, Noom D, Müller K (2024). \_igraph: Network Analysis and Visualization in R\_. <doi:10.5281/zenodo.7682609> \<[[https://doi.org/10.5281/zenodo.7682609\\\\](https://doi.org/10.5281/zenodo.7682609\\){.uri}]([https://doi.org/10.5281/zenodo.7682609\\](https://doi.org/10.5281/zenodo.7682609\){.uri}){.uri}\>, R package version 2.1.1.9005, \<[[https://CRAN.R-project.org/package=igraph\\\\](https://CRAN.R-project.org/package=igraph\\){.uri}]([https://CRAN.R-project.org/package=igraph\\](https://CRAN.R-project.org/package=igraph\){.uri}){.uri}\>.

@@

-   Pedersen T (2024). *ggraph: An Implementation of Grammar of Graphics for Graphs and Networks*. R package version 2.2.1.9000, <https://github.com/thomasp85/ggraph>, [https://ggraph.data-imaginist.com](https://ggraph.data-imaginist.com/).

------------------------------------------------------------------------

### 2024-11-18 draftintro.Rmd
visualtextdarkmodecorrections

```{=html}
<style>
body {
text-align: left;
font-family: Times;
}

h1, .h1, h2, .h2, h3, .h3 {
  margin-top: 24px;
  font-family: Times;
}

@@

-   **RQ1a:** To what extent does gender homophily occur across all co-publications of the staff in the Sociology and Political Science Departments at Radboud University?

-   **RQ1b:** To what extent do the two departments differ in their levels of gender homophily?

@@

-   **RQ2:** To what extent can gender homophily in collaboration networks be explained by preferences at the dyadic level, controlling for structural network and individual-level covariates?

>>>>

```{=html}
<style>
body {
text-align: left;
line-height: 1.5;
}

h1, .h1, h2, .h2, h3, .h3 {
  margin-top: 24px;
  line-height: 1;
}

@@

-   ***RQ1a:** To what extent does gender homophily occur across all co-publications of the staff in the Sociology and Political Science Departments at Radboud University?*

-   ***RQ1b:** To what extent do the two departments differ in their levels of gender homophily?*

@@

-   ***RQ2:** To what extent can gender homophily in collaboration networks be explained by preferences at the dyadic level, controlling for structural network and individual-level covariates?*

### 2024-11-18 drafttheory.Rmd
visualtextdarkmodecorrections

@@

>>>>

### Theory

Understanding collaboration patterns in academia is crucial for unpacking the mechanisms behind the formation of scholarly networks. One key principle that shapes these networks is **homophily**, the tendency of individuals to associate with others who are similar to them (@McPherson2001). In the context of academic collaborations, **gender homophily**—the preference for scholars to collaborate with others of the same gender—has been widely observed. Research consistently shows that men tend to collaborate more frequently with other men, and women with other women ( @Horta2022; @Kwiek2021). However, the degree and dynamics of this homophily can vary across disciplines and contexts, making it a fertile ground for investigation.

#### Hypothesis 1: Gender Homophily in Collaboration Networks

The first hypothesis centers on the presence of gender homophily across two academic departments: Political Science and Sociology at Radboud University. Based on the homophily principle, I expect scholars to prefer same-gender collaborators, leading to gendered clustering in co-authorship networks.

-   ***H1**: Across the collaboration networks of both departments, scholars are more likely to collaborate with same-gender peers than with different-gender peers, indicating the presence of gender homophily.*

This hypothesis aligns with prior studies that highlight gender homophily as a key feature of collaboration networks in academia (@Kwiek2021). However, the extent to which this tendency manifests may depend on the structural characteristics of each department, as well as broader societal norms and stereotypes.

#### Hypothesis 2: Variation in Gender Homophily Between Departments

While gender homophily is a general phenomenon, its intensity may vary across disciplines due to their differing gender dynamics and societal perceptions. Sociology, often perceived as a more gender-balanced or even female-leaning field, may exhibit weaker gender homophily compared to Political Science, which is often associated with traditionally masculine traits such as rationality and assertiveness. These stereotypes may influence collaboration patterns, with scholars in more male-dominated fields showing stronger gender homophily (@Kwiek2021).

-   ***H2**: Scholars in the Sociology Department exhibit weaker gender homophily in their collaborations than those in the Political Science Department.*

This hypothesis builds on the idea that the gender composition and cultural norms of academic disciplines shape collaboration preferences. In sociology, where gender parity is more pronounced, collaboration across genders may be more normalized, reducing homophilic tendencies.

#### Hypothesis 3: The Influence of Prestige on Gender Homophily

While homophily can drive collaboration patterns, prestige also plays a critical role. Scholars are often drawn to collaborate with those who have high academic prestige, typically measured through metrics such as publication in high-impact journals or h-index scores (@Bornmann2020). This "prestige effect" may weaken the influence of gender homophily, as the opportunity to collaborate with a prestigious scholar may outweigh same-gender preferences.

-   ***H3**: Gender homophily is less influential in collaboration decisions when the potential collaborator has a high level of academic prestige (e.g., a strong publication record in high-prestige journals).*

This hypothesis posits that prestige serves as a countervailing force to homophily. Collaboration with high-prestige individuals is a strategic choice, driven by the potential for increased visibility and career advancement, which may diminish the salience of gender in collaboration decisions.

### Theoretical Framework

The theoretical framework draws on both **homophily theory** and the concept of **preferential attachment** in academic networks. The homophily principle explains the tendency for scholars to form ties with others who share their gender, while preferential attachment highlights the allure of prestigious collaborators ( @Merton1968; @Broido2019). Together, these mechanisms shape the structure and evolution of collaboration networks.

Prior research suggests that men, particularly in male-dominated fields, are more likely to form same-gender ties, reinforcing gender inequalities in academia (@Kwiek2021). However, as women gain visibility and prestige, their collaboration opportunities expand, potentially mitigating homophily effects (@Knobloch2013). Understanding these dynamics requires a nuanced approach that accounts for both individual preferences and structural network constraints.

By addressing these hypotheses, this study contributes to the broader literature on academic collaboration and gender dynamics, offering insights into how societal norms and institutional contexts influence the formation of scholarly networks.

# prior notes:

@@

Broido, A. D., & Clauset, A. (2019). Scale-free networks are rare. *Nature communications*, *10*(1), 1017. <https://doi.org/10.1038/s41467-019-08746-5>

@Broido2019

Merton, Robert K. 1968. “The Matthew Effect in Science: The Reward and Communication Systems of Science Are Considered.” *Science* 159 (3810): 56–63.

@Merton1968

Knobloch-Westerwick, Silvia, Carroll J Glynn, and Michael Huge. 2013. “The Matilda Effect in Science Communication: An Experiment on Gender Bias in Publication Quality Perceptions and Collaboration Interest.” *Science Communication* 35 (5): 603–25.

@Knobloch2013

------------------------------------------------------------------------

### 2024-11-18 weekbacktrack.Rmd
visualtextdarkmodecorrections

| Continue description & plotting [S3]                                                              |                                                              |                                                                 | Continue description & plotting [S3]                                                       | Data & methods part [W]                                                       |                                                                            | estimate model [S5]                                                                  |
|                                                                                                   |                                                              |                                                                 |                                                                                            |                                                                               |                                                                            | goodness of fit [S6]                                                                 |

@@

| Check in with Jochem                                                                              |                                                              |                                                                 | Results [W]                                                                                | Finish Results [W]                                                            | email Jochem if delay                                                      | Rework intro [W]                                                                     |

>>>>

| Continue description & plotting [S3]                                                              |                                                              |                                                                 | Continue description & plotting [S3]                                                       | Data & methods part [W]                                                       |                                                                            |                                                                                      |

@@

| Check in with Jochem!                                                                             |                                                              |                                                                 | estimate model [S5]                                                                        | Results [W]                                                                   | email Jochem if delay                                                      | Rework intro [W]                                                                     |
|                                                                                                   |                                                              |                                                                 | goodness of fit [S6]                                                                       |                                                                               |                                                                            |                                                                                      |
